---
title: "Homework 4"
author: "Alain Kuiete"
date: "11/5/2020"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
```

## DATA EXPLORATION
```{r}
auto.insurance <- read.csv('insurance_training_data.csv')
head(auto.insurance)
#write.csv(head(auto.insurance), "head_insurance_data.csv")
```


```{r}
cost.ins <- auto.insurance
#attach(cost.ins)
```


```{r}
evaluation <- read.csv("insurance-evaluation-data.csv")
str(evaluation)
```

#### Remove $ sign on INCOME, HOME_VAL, BLUEBOOK, and OLDCLAIM

```{r}
dollar.sign <- function(a){
  a <- as.numeric(gsub("[\\$,]", "", a))
  return(a)
}
cost.ins$INCOME <- dollar.sign(cost.ins$INCOME)
cost.ins$HOME_VAL <- dollar.sign(cost.ins$HOME_VAL)
cost.ins$BLUEBOOK <-  dollar.sign(cost.ins$BLUEBOOK)
cost.ins$OLDCLAIM <- dollar.sign(cost.ins$OLDCLAIM)
summary(cost.ins)
write.csv(summary(cost.ins), "summary_data.csv")
```
### Evaluation
```{r}
evaluation$INCOME <- dollar.sign(evaluation$INCOME)
evaluation$HOME_VAL <- dollar.sign(evaluation$HOME_VAL)
evaluation$BLUEBOOK <-  dollar.sign(evaluation$BLUEBOOK)
evaluation$OLDCLAIM <- dollar.sign(evaluation$OLDCLAIM)
summary(evaluation)
```


#### Transforming to categorical variables

```{r}
cost.ins$TARGET_FLAG <- factor(cost.ins$TARGET_FLAG, labels = c("NO","YES"))
cost.ins$HOMEKIDS <- factor(cost.ins$HOMEKIDS, labels = c("HK1", "HK2", "HK3", "HK4", "HK5", "HK6"))
cost.ins$CLM_FREQ <- factor(cost.ins$CLM_FREQ, labels = c("CF1", "CF2", "CF3", "CF4", "CF5", "CF6"))
#cost.ins$KIDSDRIV <- factor(cost.ins$KIDSDRIV)
cost.ins$URBANICITY <- ifelse(cost.ins$URBANICITY=="Highly Urban/ Urban", "U","R")
cost.ins$MSTATUS <- ifelse(cost.ins$MSTATUS=="Yes", "1","0")
cost.ins$SEX <- ifelse(cost.ins$SEX=="M", "1","0")
cost.ins$CAR_USE <-  ifelse(cost.ins$CAR_USE=="Private", "1","0")
cost.ins$EDUCATION <- ifelse(cost.ins$EDUCATION=="<High School","1HS",
                             ifelse(cost.ins$EDUCATION=="z_High School","2HS",
                                    ifelse(cost.ins$EDUCATION=="Bachelors","3BA",
                                           ifelse(cost.ins$EDUCATION=="Masters","4MA","5PH"))))
#cost.ins$JOB[cost.ins$JOB==""] <- "UNEMPLOY"
summary(cost.ins)
```
```{r}
head(cost.ins)
```

### Evaluation


```{r}
#evaluation$TARGET_FLAG <- factor(evaluation$TARGET_FLAG, labels = c("NO","YES"))
evaluation$HOMEKIDS <- factor(evaluation$HOMEKIDS, labels = c("HK1", "HK2", "HK3", "HK4", "HK5", "HK6"))
evaluation$CLM_FREQ <- factor(evaluation$CLM_FREQ, labels = c("CF1", "CF2", "CF3", "CF4", "CF5", "CF6"))
#cost.ins$KIDSDRIV <- factor(cost.ins$KIDSDRIV)
evaluation$URBANICITY <- ifelse(evaluation$URBANICITY=="Highly Urban/ Urban", "U","R")
evaluation$MSTATUS <- ifelse(evaluation$MSTATUS=="Yes", "1","0")
evaluation$SEX <- ifelse(evaluation$SEX=="M", "1","0")
evaluation$CAR_USE <-  ifelse(evaluation$CAR_USE=="Private", "1","0")
evaluation$EDUCATION <- ifelse(evaluation$EDUCATION=="<High School","1HS",
                             ifelse(evaluation$EDUCATION=="z_High School","2HS",
                                    ifelse(evaluation$EDUCATION=="Bachelors","3BA",
                                           ifelse(evaluation$EDUCATION=="Masters","4MA","5PH"))))
#cost.ins$JOB[cost.ins$JOB==""] <- "UNEMPLOY"
summary(evaluation)
```



```{r}
attach(cost.ins)
str(cost.ins)
```




### Univariate Study
#### Histogram of some variables
We divide the variable in two groups to study




```{r}
library(DataExplorer)
var.group1 <-data.frame(TARGET_AMT, KIDSDRIV, AGE, YOJ, INCOME, HOME_VAL, TRAVTIME, TIF, 
                        OLDCLAIM, BLUEBOOK, MVR_PTS, CAR_AGE)
plot_histogram(var.group1, ggtheme = theme_light())
```

```{r}
ggplot(cost.ins, aes(x=EDUCATION)) + geom_bar()
```

```{r}
ggplot(cost.ins, aes(CAR_TYPE)) + geom_bar()
```

```{r}
ggplot(cost.ins, aes(x=TARGET_AMT)) + geom_histogram(binwidth = 4000)
```



```{r}
str(cost.ins)
```


```{r}
ggplot(cost.ins, aes(x=TARGET_FLAG, y=TARGET_AMT)) + geom_boxplot()
```

### Bivariate Study

## DATA PREPARETION




```{r}

```



```{r}

```



#### Looking for correlations between variables
```{r}
income.group <- data.frame(INCOME, EDUCATION, JOB, HOME_VAL, CAR_TYPE, BLUEBOOK, HOMEKIDS, RED_CAR, URBANICITY, TRAVTIME, REVOKED, CAR_USE, CLM_FREQ, OLDCLAIM, MVR_PTS)

```

```{r}
ggplot(income.group, aes(x=log(INCOME), y=log(BLUEBOOK), color = JOB)) + geom_point(position = position_jitter(), alpha=.3) + facet_grid(rows = vars(URBANICITY), cols = vars(EDUCATION))
```
There exist relative  correlation between some categorical variables. We see that in urban like in rural area, blue collar are for les than high school or more, people with bachelor and master have professional, manager and some home maker jobs, With PhD, someone is likely to be Home Maker, Doctor or Clerical




```{r}
ggplot(income.group, aes(x=MVR_PTS, y=log(TARGET_AMT), color = CLM_FREQ)) + geom_point() + facet_grid(cols = vars(REVOKED))
```

```{r}
ggplot(income.group, aes(x=log(INCOME), y=log(TARGET_AMT), color = JOB)) + geom_point() + facet_grid(rows = vars(URBANICITY), cols = vars(EDUCATION))
```


```{r}
ggplot(cost.ins, aes(x=TIF, y=YOJ, color = JOB)) + geom_point()
```


#### Missing Values in AGE , INCOME, HOME_VAL, YOJ, and CAR_AGE
Missing Values in AGE , INCOME, HOME_VAL, YOJ, and CAR_AGEThe mice package in R, helps you imputing missing values with plausible data values. 
These plausible values are drawn from a distribution specifically designed for each missing datapoint.

Here we going to form a data frame with only varibles that have missing values
```{r}
ins.missing <- data.frame(AGE, INCOME, HOME_VAL, CAR_AGE, YOJ)
```
The mice package provides a nice function md.pattern() to get a better understanding of the pattern of missing data


```{r}
ins.missing.eva <- data.frame(AGE = evaluation$AGE, INCOME = evaluation$INCOME, HOME_VAL= evaluation$HOME_VAL, CAR_AGE = evaluation$CAR_AGE, YOJ= evaluation$YOJ)
```


```{r}
library(mice)
md.pattern(ins.missing)
write.csv(md.pattern(ins.missing),"missing_data.csv")
```
There are 6448 row with no missing values. 510 missing in CAR_AGE, 464 missing in HOME_VAL, 
454 samples miss in YOJ (Year of JOB), 445 samplea miss in INCOME, and 6 samples miss in AGE



```{r}
library(VIM)
aggr_plot <- aggr(ins.missing, col=c('navyblue','red'), numbers=TRUE, sortVars=TRUE, labels=names(ins.missing), cex.axis=.7, gap=3, ylab=c("Histogram of missing data","Pattern"))
```

The plot helps us understanding that almost 90% of the samples are not missing any information, 6% are missing the CAR_AGE, and the remaining ones show other missing patterns.


```{r}
temp.ins <- mice(ins.missing,m=5,maxit=50,meth='rf',seed=123)
```

```{r}
temp.ins.eva <-mice(ins.missing.eva,m=5,maxit=50,meth='rf',seed=123)
```



```{r}
summary(temp.ins)
```

```{r}
completed.ins <- complete(temp.ins, 1)
```

```{r}
completed.ins.eva <- complete(temp.ins.eva,1)
```


```{r}
summary(completed.ins)
write.csv(summary(completed.ins), "imputed_data.csv")
```

```{r}
str(completed.ins)
```


```{r}
costomers.insurance <- cbind(cost.ins[,-c(1, 5, 7, 8, 10, 25)], completed.ins)
str(costomers.insurance)
```

```{r}
costomers.insurance.eva <- cbind(evaluation[,-c(1, 5, 7, 8, 10, 25)], completed.ins.eva)
str(costomers.insurance.eva)
```


```{r}
summary(costomers.insurance)
```









Missing Values in JOB
Detecting the missing values
```{r}
#levels(cost.ins$JOB)
#cost.ins$JOB[cost.ins$JOB==""]
```





### Split the data into train and test set
```{r}
library(caret)
set.seed(25)
inTraining <- createDataPartition(costomers.insurance$TARGET_AMT, p = .75, list = FALSE)
training <- costomers.insurance[ inTraining,]
testing  <- costomers.insurance[-inTraining,]
```

## BUILD MODELS
### Build Models with the trainset using k-10 fold
#### Multiple Linear Model
Linear regression using:
1. caret package
```{r}
fitControl <- trainControl(## 10-fold CV
  method = "repeatedcv",
  number = 3,
  ## repeated ten times
  repeats = 10)


set.seed(25)
gbmFit1 <- train(TARGET_AMT ~ ., data = training, 
                 method = "gbm", 
                 trControl = fitControl,
                 ## This last option is actually one
                 ## for gbm() that passes through
                 verbose = FALSE)
gbmFit1

```

```{r}
summary(gbmFit1)
write.csv(summary(gbmFit1), "gbmFit1_varimp.csv")
```




Selecting models
```{r}
plot(gbmFit1)

```


##### Classification

GBM model
```{r}
gbmGrid <-  expand.grid(interaction.depth = c(1, 5, 9),
                        n.trees = (1:30)*50,
                        shrinkage = 0.1,
                        n.minobsinnode = 20)

nrow(gbmGrid)

set.seed(25)
gbmFit2 <- train(TARGET_AMT ~ ., data = training,
                 method = "gbm",
                 trControl = fitControl,
                 verbose = FALSE,
                 ## Now specify the exact models
                 ## to evaluate:
                 tuneGrid = gbmGrid)
gbmFit2
```



2. Partial Lesat Square
```{r}
set.seed(25)
fitControl <- trainControl(
  method = "repeatedcv",
  classProbs = FALSE,
  repeats = 3,
  preProc = c("center", "scale"))

plsFit1 <- train(TARGET_AMT~., data=training, method = "pls", 
                 tuneLength=15, metric="RMSE", 
                 trControl=fitControl)
plsFit1
```


```{r}
plsFit1$results
write.csv(plsFit1$results, "plsFit1_result.csv")
```

```{r}
plsFit1
```


3. Using step function
```{r}
step.lmod <- lm(TARGET_AMT ~ ., data = training)
step.lmod <- step(step.lmod, trace = FALSE)
```



```{r}
summary(step.lmod)
```


```{r}

```


```{r}
predictor.importance <-  varImp(step.lmod)
predictor.importance
write.csv(predictor.importance, "predictor_imp.csv")
```
Most important predictors are UBANICITY, MVR_PTS, MSTATUS, CAR_USE, CAR_TYPE, TIF, TRAVTIME, KIDSDRIVKD, REVOKED, CAR_AGE, PARENT1, INCOME, AGE, HOME_VAL, 




4. rlm from MASS package
We use the Huber weigth parameter
```{r}
library(MASS)
rr.huber <- rlm(TARGET_AMT ~ KIDSDRIV + PARENT1 + MSTATUS + JOB + 
    TRAVTIME + CAR_USE + TIF + CAR_TYPE + REVOKED + MVR_PTS + 
    URBANICITY + AGE + INCOME+ HOME_VAL + CAR_AGE, data = training)
rr.huber
```

```{r}
sqrt(mean(rr.huber$residuals^2)) 
```
```{r}
summary(rr.huber)
```


```{r}
library(leaps)
sub.models <- regsubsets(TARGET_AMT~., data = training, method = "exhaustive", nvmax = 32)
rs <- summary(sub.models)
rs$which
```


Examine model further - list outputs available with names() :
```{r}
best.summary <- summary(sub.models)
names(best.summary)
```

Explore model with lowest RSS and highest R-squared:
```{r}
which.min(best.summary$rss)
which.max(best.summary$adjr2)
which.min(best.summary$bic)
```


#### Binary Logistic Regression
### Split the data into train and test set

```{r}
library(caret)
set.seed(25)
inTraining <- createDataPartition(costomers.insurance$TARGET_AMT, p = .75, list = FALSE)
training <- costomers.insurance[ inTraining,-2]
testing  <- costomers.insurance[-inTraining,-2]
```

1. glm logit and probit
```{r}
logitFit <- glm(TARGET_FLAG ~ KIDSDRIV + PARENT1 + MSTATUS + JOB + 
    TRAVTIME + CAR_USE + TIF + CAR_TYPE + REVOKED + MVR_PTS + 
    URBANICITY + AGE + log(INCOME) + HOME_VAL + CAR_AGE, family = binomial(link = "logit"), data = cost.ins)
logitFit
```


```{r}
length(testing$KIDSDRIV)
```


```{r}
library(caret)
pred <- predict(logitFit, newdata = testing[,-1], type = "response")
predicted <- as.factor(ifelse(pred>=.5,"YES","NO"))
cm <- confusionMatrix(data = predicted, testing$TARGET_FLAG, positive = 'YES')
write.csv(cm$table, "confusion_matrix_table.csv")
write.csv(cm$byClass, "confusion_matrix_metrics.csv")
cm$byClass
```


```{r}
metric.logit <- logitFit
```




Selecting models
2. GBM with tuning parameters
```{r}

gbmGrid <-  expand.grid(interaction.depth = c(1, 5, 9),
                        n.trees = (1:30)*50,
                        shrinkage = 0.1,
                        n.minobsinnode = 20)

nrow(gbmGrid)

fitControl <- trainControl(method = "repeatedcv",
                           number = 5,
                           repeats = 3,
                           ## Estimate class probabilities
                           classProbs = TRUE,
                           ## Evaluate performance using 
                           ## the following function
                           summaryFunction = twoClassSummary)

set.seed(25)
gbmFit3 <- train(TARGET_FLAG~ ., data = training, 
                 method = "gbm", 
                 trControl = fitControl, 
                 verbose = FALSE, 
                 tuneGrid = gbmGrid,
                 ## Specify which metric to optimize
                 metric = "ROC")
gbmFit3
```


```{r}
whichTwoPct <- tolerance(gbmFit3$results, metric = "ROC", 
                         tol = 2, maximize = FALSE)  
cat("best model within 2 pct of best:\n")
```

```{r}
best.model <- gbmFit3$results[whichTwoPct,1:6]
write.csv(best.model, "best.model.csv")
```


Between Models
Support Vector Machine is fit to the training set
```{r}
set.seed(25)
svmFit <- train(TARGET_FLAG ~ ., data = training, 
                 method = "svmRadial", 
                 trControl = fitControl, 
                 preProc = c("center", "scale"),
                 tuneLength = 8,
                 metric = "ROC")
svmFit                 
```


Discriminant Analysis is fit 
```{r}
set.seed(25)
rdaFit <- train(TARGET_FLAG ~ ., data = training, 
                 method = "rda", 
                 trControl = fitControl, 
                 tuneLength = 4,
                 metric = "ROC")
rdaFit             
```

Statistical Statements on the models perfomances

Resampling Results
```{r}
resamps <- resamples(list(GBM = gbmFit3,
                          SVM = svmFit,
                          RDA = rdaFit))
resamps
```

Visualizing the resampling distribution
```{r}
theme1 <- trellis.par.get()
theme1$plot.symbol$col = rgb(.2, .2, .2, .4)
theme1$plot.symbol$pch = 16
theme1$plot.line$col = rgb(1, 0, 0, .7)
theme1$plot.line$lwd <- 2
trellis.par.set(theme1)
bwplot(resamps, layout = c(3, 1))
```

```{r}
trellis.par.set(caretTheme())
dotplot(resamps, metric = "ROC")
```

```{r}
trellis.par.set(theme1)
xyplot(resamps, what = "BlandAltman")
```



```{r}
splom(resamps)
```

Since models are fit on the same versions of the training data, it makes sense to make inferences on the differences between models. In this way we reduce the within-resample correlation that may exist. We can compute the differences, then use a simple t-test to evaluate the null hypothesis that there is no difference between models.

```{r}
difValues <- diff(resamps)
difValues
```

## Using the whole training set
```{r}

gbmGrid <-  expand.grid(interaction.depth = c(1, 5, 9),
                        n.trees = (1:30)*50,
                        shrinkage = 0.1,
                        n.minobsinnode = 20)

nrow(gbmGrid)

fitControl <- trainControl(method = "repeatedcv",
                           number = 10,
                           repeats = 10,
                           ## Estimate class probabilities
                           classProbs = TRUE,
                           ## Evaluate performance using 
                           ## the following function
                           summaryFunction = twoClassSummary)

set.seed(25)
gbmFit4 <- train(TARGET_FLAG~ ., data = costomers.insurance[,-2], 
                 method = "gbm", 
                 trControl = fitControl, 
                 verbose = FALSE, 
                 tuneGrid = gbmGrid,
                 ## Specify which metric to optimize
                 metric = "ROC")

```

```{r}
pred
```


#### Metrics
```{r}
pred <- predict(rdaFit, newdata = testing[,-1], type = "raw")
#predicted <- as.factor(ifelse(pred>=.5,"YES","NO"))
cm <- confusionMatrix(data = pred, testing$TARGET_FLAG, positive = 'YES')
write.csv(cm$table, "confusion_matrix_table6.csv")
write.csv(cm$byClass, "confusion_matrix_metrics6.csv")
cm$byClass
```







## Making prediction




```{r}
evaluated <- predict(gbmFit4, newdata = evaluation, type ="prob")
head(evaluated)
write.csv(evaluated, "evaluated_insurance.csv")
```


```{r}
predicted.ins.eva <-  predict(gbmFit4, newdata = costomers.insurance.eva)
summary(predicted.ins.eva)
```


```{r}
to.predict <- read.csv("insurance-evaluation-data.csv")
predict.insur <- cbind(to.predict[,-2], TARGET_FLAG = predicted.ins.eva )
head(predict.insur)
```



### Amount of Insurance

```{r}
amount <- predict(gbmFit1, newdata =costomers.insurance.eva )
target.amount <- ifelse(prediction.insurance$TARGET_FLAG=="NO",0 ,amount)
predict.insur <- cbind(predict.insur, TARGET_AMT = target.amount )
head(predict.insur)

```

```{r}
write.csv(predict.insur, "insurance_evaluation.csv")
```


